---
title: "Competitive debate record"
author: "James Laird-Smith"
date: 'Last updated: `r format(Sys.Date(), "%Y-%m-%d")`'
lastmod: '`r format(Sys.Date(), "%Y-%m-%d")`'
summary: "My debate CV"
draft: false
output:
  blogdown::html_page:
    toc: false

---
<style>
.article-container {
  max-width: 1200px;
}

.article-date {
  display:none;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(googlesheets4)
library(reactable)
library(janitor)
library(emo)
```


```{r include=FALSE}
# cv_entries_judging_from_sheet <-
#   read_sheet("1aWslvXXbs70PLCkEnvtjAdAGxoOFiYFEK2y5IiJg9Zk",
#              sheet = "Combined")
# 
# cv_entries_judging_from_sheet %>%
#   write_csv("judging_entries.csv")
```


```{r include=FALSE}
cv_entries_judging_raw <- read_csv("judging_entries.csv")
```


```{r include=FALSE}
cv_entries_judging <- 
  cv_entries_judging_raw %>% 
  clean_names() %>% 
  mutate(date_first_day = as.Date(date_first_day),
         reached = str_remove(reached, "Broke to ")) %>%
  mutate(reached = if_else(reached == "Didn't break",
                           NA_character_,
                           reached)) %>% 
  mutate(broke = if_else(broke,
                         as.character(emo::ji("check")),
                         as.character(emo::ji("multiplication")))) %>% 
  mutate(competition_type = if_else(competition_type == "IV",
                                    "Universities",
                                    competition_type)) %>% 
  arrange(rank) %>% 
  select(competition,
         year,
         role,
         broke,
         competition_type,
         reached,
         date_first_day) %>% 
  mutate(rn = row_number()) %>% 
  select(rn, everything())
```

For a big part of my life, I have involved myself in competitive debating.
This is different from ordinary debate in that participants rarely 
know the topics in advance and cannot choose the position from which they argue. 
The winners and losers of a competitive debate are not chosen by an audience, 
but rather by a panel of other competitive debaters. 

Much of competitive debating is centered around competitions. These take place 
all over the world and involve many rounds of debate happening over a number of
days. Because competitive debate is still a student-centered 
activity, most of these competitions are hosted by university debate societies.

The following table is a summary of some of my performances, both as a speaker
and a judge, at various debate competitions.

Each of the rows represents a different competition. To "break" at a debate
competition is a good thing: it means to progress through to the elimination 
rounds. In these later stages teams compete to knock each another out and 
so fewer panels of judges are required. As such, the quality of a speaker or
judge is indicated by how far into a competition they progress.<br>
<br>

```{r echo=FALSE}
colum_specs <- list(
  year = colDef(name = "Year", align = "left", maxWidth = 70),
  role = colDef(name = "Role", align = "left", maxWidth = 95),
  competition = colDef(name = "Competition name"),
  competition_type = colDef(name = "Competition type", minWidth = 120),
  broke = colDef(name = "Broke", maxWidth = 70),
  reached = colDef(name = "Furthest stage reached", na = "-", minWidth = 120),
  rn = colDef(name = "#", maxWidth = 45, align = "left"),
  date_first_day = colDef(name = "Start date")
)

reactable(cv_entries_judging,
          pagination = FALSE,
          columns = colum_specs)
```