---
title: "Competitive debate record"
author: "James Laird-Smith"
date: 'Last updated: `r format(Sys.Date(), "%Y-%m-%d")`'
summary: "My debate CV"
draft: false
output:
  blogdown::html_page:
    toc: false

---
<style>
.article-container {
  max-width: 1200px;
}

.article-date {
  display:none;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(googlesheets4)
library(reactable)
library(janitor)
library(emo)
```


```{r include=FALSE}
# cv_entries_judging_from_sheet <-
#   read_sheet("1aWslvXXbs70PLCkEnvtjAdAGxoOFiYFEK2y5IiJg9Zk",
#              sheet = "Judging")
# 
# cv_entries_judging_from_sheet %>%
#   write_csv("judging_entries.csv")
```


```{r include=FALSE}
cv_entries_judging_raw <- read_csv("judging_entries.csv")
```


```{r include=FALSE}
cv_entries_judging <- 
  cv_entries_judging_raw %>% 
  clean_names() %>% 
  mutate(date_first_day = as.Date(date_first_day),
         reached = str_remove(reached, "Broke to ")) %>%
  mutate(reached = if_else(reached == "Didn't break",
                           NA_character_,
                           reached)) %>% 
  mutate(broke = if_else(broke,
                         as.character(emo::ji("check")),
                         as.character(emo::ji("multiplication")))) %>% 
  mutate(competition_type = if_else(competition_type == "IV",
                                    "Intervarsity",
                                    competition_type)) %>% 
  select(competition,
         year,
         competition_type,
         broke,
         reached,
         date_first_day) %>% 
  mutate(rn = row_number()) %>% 
  select(rn, everything())
```

For a big part of my life, I have involved myself in competitive debating.
This is a different activity from ordinary debate in that partisipants rarely 
know the topics in advance and cannot choose the position from which they argue. 
The winners and losers of a competitive debate are not chosen by an audience, 
but rather by a panel of other competitve debaters. 

Much of competitve debating is centered around competitions which take place 
all over the world and involve many rounds of debate happening over a number of
days. As competitive debate is still a student-centered vocation, most of the
competitions are hosted by university debate socioties.

The following table is a summary of some of my performances as a judge at
various debate competitions.

Each of the rows represents a different competition. To "break" at a debate
competition means to progress through to the elimination rounds. In the later
stages the teams compete to eliminate each another and so fewer panels of
judges are required as the competition moves towards the final. As such, the
further into the competition a judge progresses is taken to be an indicator 
of their judging ability.

```{r echo=FALSE}
colum_specs <- list(
  year = colDef(name = "Year", align = "left", maxWidth = 70),
  competition = colDef(name = "Competition name"),
  competition_type = colDef(name = "Competition type", minWidth = 120),
  broke = colDef(name = "Broke", maxWidth = 70),
  reached = colDef(name = "Furthest stage reached", na = "-", minWidth = 120),
  rn = colDef(name = "#", maxWidth = 45, align = "left"),
  date_first_day = colDef(name = "Start date")
)

reactable(cv_entries_judging,
          pagination = FALSE,
          columns = colum_specs)
```