---
title: James' unofficial guide to attending WUDC Cape Town 2019
author: "James Laird-Smith"
date: '2018-11-02'
slug: james-guide-to-attending-ct-worlds
categories: []
tags: []
header:
  caption: ''
  image: ''
summary: " "
draft: true
bibliography: clothing-search.bib
link-citations: true
csl: cell-numeric-superscript.csl
---

```{r setup, include=FALSE}
library(tidyverse)
library(highcharter)
library(widgetframe)
library(r2d3)
library(lubridate)
library(htmltools)
```

The World Universities Debating Championships (WUDC) this year is being hosted at the University of Cape Town (UCT), where I was enrolled for both my undergraduate and postgraduate degrees. For those, like me, who are attending the competition, I thought it would be useful if I put together some thoughts on what it will be like once you get there. These thoughts are catalogued below. Please note that, although I did attend the university for a good long time, this is a decidedly unofficial guide designed for supplementary information only. Please defer to the Organising Committee (Org Comm) for all the official information regarding the competition.

## Arriving

UCT has a few campuses. You only need to worry about one of them and that is the **Lower Campus**. This is where the university's residences are located and where all the participants will be staying.

## Getting around

Uber works in Cape Town the same way it does in many cities around the world. This makes getting around relatively easy.


## What to wear






```{r, echo=FALSE}
htmltools::includeHTML("ct_rain_files/ct_rain.html")
```


```{r, echo=FALSE}
htmltools::includeHTML("ct_rain_files/ct_dam_levels.html")
```

```{r eval=FALSE, include=FALSE}
r2d3(data = dam_levels_data, script = "damLevelBars.js",
     container = "div",
     width =  "100%",
     sizing = sizingPolicy(defaultWidth = "100%"),
     css = "standardChartStyle.css"
     )#
```


